### 2022-08-11 fig4.tex

\begin{figure*}%[ht]
\begin{center}
\includegraphics[angle=0,width=4.0cm]{figures/thumbnails/SN2005ey.png}
\includegraphics[angle=0,width=4.0cm]{figures/thumbnails/SN2005go.png}
\includegraphics[angle=0,width=4.0cm]{figures/thumbnails/SN2005gs.png}
\includegraphics[angle=0,width=4.0cm]{figures/thumbnails/SN2005ja.png}
\includegraphics[angle=0,width=4.0cm]{figures/thumbnails/SN2005jc.png}
\includegraphics[angle=0,width=4.0cm]{figures/thumbnails/SN2005jd.png}
\includegraphics[angle=0,width=4.0cm]{figures/thumbnails/SN2005jn.png}
\includegraphics[angle=0,width=4.0cm]{figures/thumbnails/SN2006fk.png}
\includegraphics[angle=0,width=4.0cm]{figures/thumbnails/SN2006gn.png}
\includegraphics[angle=0,width=4.0cm]{figures/thumbnails/SN2006hf.png}
\includegraphics[angle=0,width=4.0cm]{figures/thumbnails/SN2006hk.png}
\includegraphics[angle=0,width=4.0cm]{figures/thumbnails/SN2006mn.png}
\includegraphics[angle=0,width=4.0cm]{figures/thumbnails/SN2006nj.png}
\includegraphics[angle=0,width=4.0cm]{figures/thumbnails/SN2006nu.png}
\includegraphics[angle=0,width=4.0cm]{figures/thumbnails/SN2006nw.png}
\includegraphics[angle=0,width=4.0cm]{figures/thumbnails/SN2006oi.png}
\includegraphics[angle=0,width=4.0cm]{figures/thumbnails/SN2006pi.png}
\includegraphics[angle=0,width=4.0cm]{figures/thumbnails/SN2007ix.png}
\includegraphics[angle=0,width=4.0cm]{figures/thumbnails/SN2007je.png}
\includegraphics[angle=0,width=4.0cm]{figures/thumbnails/SN2007jj.png}
\includegraphics[angle=0,width=4.0cm]{figures/thumbnails/SN2007jo.png}
\includegraphics[angle=0,width=4.0cm]{figures/thumbnails/SN2007lw.png}
\includegraphics[angle=0,width=4.0cm]{figures/thumbnails/SN2007mk.png}
\includegraphics[angle=0,width=4.0cm]{figures/thumbnails/SN2007ml.png}
\includegraphics[angle=0,width=4.0cm]{figures/thumbnails/SN2007nd.png}
\includegraphics[angle=0,width=4.0cm]{figures/thumbnails/SN2007pj.png}
\includegraphics[angle=0,width=4.0cm]{figures/thumbnails/SN2007qp.png}
\includegraphics[angle=0,width=4.0cm]{figures/thumbnails/SN2007qy.png}
\includegraphics[angle=0,width=4.0cm]{figures/thumbnails/SN2007rl.png}
\includegraphics[angle=0,width=4.0cm]{figures/thumbnails/SN2007rm.png}
\includegraphics[angle=0,width=4.0cm]{figures/thumbnails/SN2007rq.png}
\includegraphics[angle=0,width=4.0cm]{figures/thumbnails/SN2007se.png}
\end{center}
\end{figure*}